---
title: "mod12_project_markdown"
author: "D.Hannagan"
date: "2024-03-29"
output: pdf_document
---

#Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing #*Cmd+Option+I*.

#CANDIDA AURIS SUBSTRATE ANALYSIS

1)Need to import lab Candida experimental data:
```{r}
sth<- read.csv("~/Desktop/BioDatAnalysis/cauris72.csv")
View(sth)
```



2)Need to view data and the summary of experimental:
```{r}
head(sth)
summary(sth)
```




3) Need to id dimensions:
```{r}
ncol(sth)
nrow(sth) 
count(sth)
dim(sth)

```





START ANALYSIS BELOW AFTER IMPORTATION
MULTI-VARIANT ANALYSIS-
installed the following:
```{r}
library(vegan)
library(stringr)
library(pROC)
```



#re-naming csv, first line did not work:
```{r}
dtc<- read.csv("~/Desktop/BioDatAnalysis/cauris72.csv")

```


#to view:
```{r}
head(dtc)
summary(dtc)
```




to convert column and row names to lower case:
```{r}
names(dtc) <- tolower(names(dtc))
```




to plot original data:
```{r}
plot(dtc) #margins too large, need to filter out

```



check progress:
```{r}
head(auris)
```




rename columns:
```{r}
names(auris)<-c('isolate','dglu', 'dgal','dmal','sucr','dlac', 'lara', 'draf', 'dxyl', 'glyc', 'mann', 'malt', 'dode', 'dulc')
```



check progress:
```{r}
head(auris)

```



#to create a backup dataset:
```{r}
names(backup)<-backup$Substrates
backup$SpeciesIsolate
backup<-backup[which(names(backup)!="Substrates")]
backup<-backup[-c(1:3)]
splitcols<-backup[1:3]
```



#puts two differnt columns together:
```{r}
nwq<-cbind(splitcols,backup)

```


# explore data-filter out columns i dont want:
```{r}
apply(dat[,use.cols], 2, range)

```



#to rename rows:
```{r}
rownames(auris)<-c('aur1','aur2','aur3','aur4', 'aur5', 'aur6','aur7', 'aur8','aur9','aur10', 'dha1', 'dha2', 'dha3', 'hae1', 'hae2')
```




              
#made a vector-did not use this:
```{r}
crd<- c('aur','aur','aur','aur', 'aur', 'aur','aur', 'aur','aur','aur', 'dha', 'dha', 'dha', 'hae', 'hae')
```




#to check progress:
```{r}
head(auris)
```


#to change margins:
```{r}
par(mar=c(2,2,2,2))
```



#black and white scatterplot matrix:
```{r}
pairs(auris[,2:14],lower.panel=NULL)
```



#scatter plot matrix:
```{r}
mycols <- c("red", "green", "blue")
```



#this works to make scatterplot matrix in color:
```{r}
pairs(auris[,2:14], pch=19, cex=0.5, col=c('red','red','red','red', 'red', 'red','red', 'red','red','red', 'green', 'green', 'green', 'blue', 'blue'), lower.panel=NULL)
my_cols <- c("red", "blue", "green3")  
pairs(auris[,2:4], pch = 19,  cex = 0.5,
      col = my_cols[auris$isolate],
      lower.panel=NULL)
```





#to view plot of d galactose and d xylose:
```{r}
plot(auris$dgal)
plot(auris$dxyl)
```



#t test of d xylose:
```{r}
t.test(auris$dxyl~ifelse(auris$isolate == 'hae', 'hae', 'other'))
```


#t test of d galactose:
```{r}
t.test(auris$dgal~ifelse(auris$isolate == 'aur', 'aur', 'other'))
```



#to view and check progress:
```{r}
view(auris)
```



# Need to answer question: "Are auris and haemulonii really different?"
```{r}
mod1 <-adonis2(auris[,2:14]~auris$isolate, method="euclidean")
```



 #IF have variable=outcomes means overfitting data (not good), need to check:
```{r}
print(mod1)
view(mod1)
```


#to view mod 1:
```{r}
mod1
```


# NOTES: So far model does predict =results are highly sig bc Pr (F) 
#what is the prb that this result is by chance=0.001 (want a small number) 
#want it to be lower than 0.5
#R2=correlation coeff squared- want it to be high=predicts % variablity 
#sumofSq (variability in dataset)=calculates R2
#Sum of Sq # =how much residual error after model ran
#residuals= variability in data not predicted by model
#high F=less likely data is random
#output is categorical


#create citation for R Studio:
```{r}
citation()
```








